---
title: Descriptive Statistics for Tidy Tuesday Alt Text Project
author: "Silvia P. Canelón, Elizabeth Hare"
date: "`r Sys.Date()`"
output: html_document
---

# Summary: How many TidyTuesday posts have alt text?



``` {r, include=FALSE}
### don't display chunks in output doc by default
knitr::opts_chunk$set(echo=FALSE)
knitr::opts_chunk$set(results='hide')
knitr::opts_chunk$set(fig_path=".")
library(knitr)
library(gtsummary)
library(TidyTuesdayAltText)
library(qdap) # word_count, character_count
library(lubridate) # for extracting week of the year
library(stringr) # contains str_pad function to make all weeks 2-digits to concatenate with year
### for bar plot
library(ggplot2)

### Access data from the four years
data(package="TidyTuesdayAltText")
data(ttTweets2018, ttTweets2019, ttTweets2020, ttTweets2021)
### concatenate year data.frames into one
ttTweets4y <- rbind(ttTweets2018, ttTweets2019, ttTweets2020, ttTweets2021)
table(ttTweets4y$Year)

### create factor variable describing alt text based on the four categories
### 1. Customized alt text written by author
### 2, Field contains default text: "Image"
### 3. Scraping error - Tweet is unavailable (deleted or private)
### 4. NA Unavailable due to non-static media


ttTweets4y$AltTextCat <- ifelse(is.na(ttTweets4y$AltText), "UnavailableNonStatic",
                         ifelse(ttTweets4y$AltText == "Image", "DefaultDescription",
                         ifelse(ttTweets4y$AltText == "ScrapingError", "ScrapingError",
                                "DescriptiveText")))
table(ttTweets4y$AltTextCat, useNA="ifany")
### convert AltTextCat to factor for easier analysis
ttTweets4y$AltTextCat <- as.factor(ttTweets4y$AltTextCat)


### make df with only image descriptions = "alt text tweets"
att <- ttTweets4y[ttTweets4y$AltTextCat == "DescriptiveText", ]
dim(att)

### cound words using qdap package
### qdap::word_count, parameters digit.remove=FALSE because we expect numerals
att$wordCount <- word_count(att$AltText, digit.remove=FALSE)
summary(att$wordCount)

att$charCount <- character_count(att$AltText)
summary(att$charCount)
```


Silvia scraped 
`r nrow(ttTweets4y)` TidyTuesday Tweets posted between 2018 and 2021. I 
concatenated them into one data set. The four categories for the alt text field were

1. Default Description - Twitter fills in "Image"
2. Descriptive Text - the user supplied alt-text
3. Scraping Error - the alt text could not be retrieved
4. Alt-text unavailable due to non-static content



``` {r, results="asasis"}
tbl_cross(ttTweets4y, row=Year, col=AltTextCat,
          percent="row")
```

# Word and character counts

## Words

Words were counted using the word_count function in the qdap R package, with
the option to include digits set to TRUE because numbers have important meaning
in describing graphs. The median word count for alt text descriptions was 
`r median(att$charCount)`, with a range of 
`r min(att$wordCount)` to 
`r max(att$wordCount)`. 
The following table presents these statistics by year.


``` {r, results="asis"}
### names(att)
tbl_summary(att, by=Year,
            include=c(Year, wordCount),
            type=all_continuous() ~ "continuous",
            statistic=all_continuous() ~ c("{median}, {min}-{max}"))

```

## Characters

The maximum character count for Twitter alt-text is 420. The median character count
for all years is 
`r median(att$charCount)`, with a range of 
`r min(att$charCount)` to
`r max(att$charCount)`. 
The following table provides these statistics by year:

```{r, results="asis"}
tbl_summary(att, by=Year,
               include=c(Year, wordCount),
               type=all_continuous() ~ "continuous",
            statistic=all_continuous() ~ c("{median}, {min}-{max}"))

```


``` {r, include=FALSE}
save(ttTweets4y, file="ttTweets4y.Rda")
save(att, file="altTextTweets.Rda")
```
``` {r, include=FALSE}
### Coming back to this analysis to answer two questions
### when is the best time (highest visibility) to launch the project on Twitter?
### how many TidyTuesdays per week based on early 2021 data?
names(ttTweets4y)
class(ttTweets4y$TweetDate)
class(ttTweets4y$Year)
table(ttTweets4y$Year)
### Let's limit ourselves to 2020-2021, since the numbers seem to be increasing
ttTweetsRecent <- ttTweets4y[ttTweets4y$Year == "2020" | ttTweets4y$Year == "2021", ]
dim(ttTweetsRecent)

### use lubridate's week function to assign "week of the year" to TweetDate
ttTweetsRecent$wk <- week(ttTweetsRecent$TweetDate)

### most recent to find out how many/week
y21 <- ttTweetsRecent[ttTweetsRecent$Year == 2021, ]

### we will also want a variable in ttTweetsRecent that concatenates year and week
### pad week with 0s using package stringr's str_pad function
ttTweetsRecent$yrwk <- paste(ttTweetsRecent$Year, str_pad(ttTweetsRecent$wk, width=2,
                             side="left", pad="0"), sep="")
head(ttTweetsRecent$yrwk)

### frequency table for first part of 2021 to be kabled later
t1 <- table(y21$wk)
t1 <- data.frame(t1)
names(t1)
colnames(t1) <- c("Week", "Count")
```

# Tweets Per Week

As estimates of how many people we can get to participate in a week, let's
look at 2021 so far.

``` {r, results="asis"}
kable(t1)
```

To get an idea of the best time of year, we look at the last 13 months of data, a bar chart 
for the eyeball people and a table for me.


``` {r, results="asis"}
ggplot(ttTweetsRecent, aes(x=yrwk)) +
    geom_bar(stat="count") +
    theme(axis.text.x=element_text(angle=90, hjust=1)) +
    labs(x="Year/Week", y="Count",
    title="Total TidyTuesdaySubmissions by Week in 2020-2021")

tbl_cross(ttTweetsRecent, row=wk, col=Year)
```
